---
title: "Getting started with eco-evo model"
output: html_notebook
---

I have never used R Notebooks before, nor have I ever built a model for publication. Let's try it!

The goal here is to sketch out the main ideas that I have for the eco-evolutionary model that Holly and I proposed as a part of my NSF postdoc application. 

Here is what I proposed to do:

"I will examine how trees shift their portfolio strategies under simulated drought scenarios of varying frequency, intensity, and duration. In this stage-structured model, individual trees will be assigned fungal portfolio strategies. These strategies will be defined by two primary traits: the number of fungi a tree associates with (q) [this was theta originally, but copy-paste made it q, which is fine with me], a measure of specificity, and the degree to which those fungi improve tree performance, on average, in the current environment (s) [originally sigma]. Both of these traits will influence the demographic rates of the trees and their reproductive outputs. New portfolio strategies will emerge occasionally via mutation. I will use this model to explore the adaptive dynamics of this tree-fungal mutualism, identifying optimal portfolio strategies under each environment, and determining which may represent evolutionarily stable strategies for trees"

So. To make this happen I will need a population of trees. This population has a couple of key properties:

A distribution of q values (from low specificity to high specificity)

A distribution of s values in environment A (from low benefit to high benefit)

A distribution of s values in environment B (from low to high benefit)

I will also eventually need a "population" of drought scenarios. These will vary in 

frequency

intensity

duration

With some reasonable probabilities associated with each state.

I'll start with the trees to try to get a simulated population going. Then I will try two environmental conditions. Then I'll see if I can get the populations to evolve via simulation. Only THEN will I layer in more nuanced drought scenarios.

Each tree has a q value and an s value.

Let's say q can range from 1 (literally just one fungus can grow with this tree) to 100 (tree is promiscuous)

We're going to have a forest of 100 trees.

```{r}
# Defining q (specificity distribution)
forest_size = 100
mean_specificity = 50
sd_specificity = 20

q = rnorm(forest_size, mean = mean_specificity, sd = sd_specificity)
hist(q)

```

Are all these fungi equally useful? Probably not. Let's treat the benefit axis (s) as a log growth response for now, so it lines up well with empirical results. A log growth response is:

ln(mass of tree with fungi/mean mass of trees without fungi)

Conveniently, this metric produces negative values when a fungus is parasitic, and positive values when the fungus is helpful. (I don't know if this will prove unwieldy in a modeling context. Let's see how it goes.)

In realistic circumstances, it would be surprising if a fungus reduced a plant's growth by more than 50% without just killing it. (Perhaps I should work in a "the plant died" option down the line.) And the greatest growth IMPROVEMENT I've ever seen in a seedling was 5X, with Arbutus menziesii seedligs and Thelephora terrestris in a very low-N artificial soil after about a year of growth. (Note: Eventually I should tie these values to a lit review. But, for now, we're building this model based on my instincts.)

```{r}
# Defining s (benefit distribution)

mean_benefit_factor = 2 # let's say that plants with fungi generally grow twice as big as plants without
sd_benefit_factor = mean_benefit_factor*.2 # just a guess that most plants get within +/- 20% of the mean benefit

s = rnorm(forest_size, mean = mean_benefit_factor, sd = sd_benefit_factor)


hist(s)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

